MODULATING TITLE LAYOUT

---
  title: "Shiny HTML Doc"
author: "theforestecologist"
date: "Apr 14, 2017"
output:  html_document
runtime: shiny
---
  
  <style type="text/css">
  
  h1.title {
    font-size: 38px;
    color: DarkRed;
    text-align: center;
  }
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>
  
----------------------------------------------------------------------------

INSERTING CLICK BUTTON

<details><summary>Click here</summary>
Some text

```{r code}
# even code
print("Hello world!")
```
</details>

----------------------------------------------------------------------------

EXTRACTING TWEETS

To learn more about text mining in R we recommend this book.

In general, we can extract data directly from twitter using the \emph{rtweet} package. However, in this case, a group has already compiled data for us and made it available at http://www.trumptwitterarchive.com.

url <- 'http://www.trumptwitterarchive.com/data/realdonaldtrump/%s.json'

trump_tweets <- map(2009:2017, ~sprintf(url, .x)) %>%
  map_df(jsonlite::fromJSON, simplifyDataFrame = TRUE) %>%
  filter(!is_retweet & !str_detect(text, '^"')) %>%
  mutate(created_at = parse_date_time(created_at, orders = "a b! d! H!:M!:S! z!* Y!", tz="EST")) 


---------------------------------------------------------------------------- 

FUNCTION extract()

```{r Extracting tweets}
df <- trump_tweets %>% 
  extract(source, "source", "Twitter for (.*)") %>%
  count(source) 
```

----------------------------------------------------------------------------

REMOVING A STRING OR JUST A CHARACTER FROM CORPUS IN JUST 2 CODE LINES

toSpace <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
corpus <- tm_map(corpus, toSpace, "[']")

----------------------------------------------------------------------------

# Converting back to character.
expanded <- expanded %>% 
  mutate(text = convert.tm.to.character(corpus)) 
  
----------------------------------------------------------------------------
  