---
title: "Popularity Prediction"
subtitle: "New York Times Blog Articles"
author: "Philippe Lambot"
date: "May 15, 2020"

output: 
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    highlight: espresso
    df_print: paged
    theme: readable
    code_folding: hide
---

```{r Setup}
# Avoiding messages and warnings: anyway, they have already been dealt with.  
knitr::opts_chunk$set(echo = T, message = F, warning = F)

# Regulating figure layout.
knitr::opts_chunk$set(out.width = "60%", fig.align = "center")

# Facilitating table layout in HTML.
options(knitr.table.format = "html")

# Output language
invisible(Sys.setlocale("LC_ALL", "C"))
```

<style type = "text/css">                  # Title colors
h1 {color: #08457E}
h2 {color: #08457E}
h3 {color: #08457E}
h4 {color: #08457E}
</style>

<style type = "text/css">                  # No bullets in TOC
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
</style>

<style type = "text/css">                  # Font size
<font size = "3">
</style>

<style type = "text/css">                  # Double text justification
body {
text-align: justify}
</style>

```{r Cleaning up workspace and downloading packages}
# Cleaning up workspace for RAM management.
invisible(if(!is.null(dev.list())) dev.off())
rm(list=ls())
cat("\014")

library(tidyverse)
library(wordcloud2)
df <- data.frame(mot = c("toto", "%", "$"), freq = c(250, 100, 50))
# set.seed(1)
wordcloud2(df, size = 1, minSize = 0, gridSize =  0,
           fontFamily = 'Segoe UI', fontWeight = 'bold',
           color = 'random-dark', backgroundColor = "blue",
           minRotation = -pi/4, maxRotation = pi/4, shuffle = TRUE,
           rotateRatio = 0.4, shape = 'circle', ellipticity = 0.65)
```

-----------------------------------------------------------------

<br>

# Executive summary {.unnumbered}

# DATA PREPARATION

## Pourquoi pas ?

### Et si !

```{r cars}
df <- summary(cars) 
str(df)

# Calculating quartiles.
s <- summary(cars)

df <- data.frame(intervals = c("Min", "1st", "Median", "Mean", "3rd", "Max"), 
                 values = matrix(s, nrow = 6, ncol = 2), 
                 stringsAsFactors = FALSE)

# Removing the mean from the data frame to only keep quartiles, minimum and maximum.
row.to.remove <- which(df$intervals %in% "Mean")
df <- df[- row.to.remove,] 

# Enlarging range to be sure of including minimum and maximum.
v <- df$values
v[1] <- v[1]

# Rounding quartiles to upgrade clarity in graph and table.
for (i in 1:(length(v) - 2))  {
  v[i + 1] <- v[i + 1]
}

library(tidyverse)
df <- df %>% mutate(values = v)
df

```

## Including Plots

You can also embed plots, for example:

```{r pressure}
invisible(if(!is.null(dev.list())) dev.off())
cat("\014")
library(wordcloud2)
df <- data.frame(mot = c("toto", "%", "$"), freq = c(250, 100, 50))
# set.seed(1)
wordcloud2(df, size = 1, minSize = 0, gridSize =  0,
           fontFamily = 'Segoe UI', fontWeight = 'bold',
           color = 'random-dark', backgroundColor = "blue",
           minRotation = -pi/4, maxRotation = pi/4, shuffle = TRUE,
           rotateRatio = 0.4, shape = 'circle', ellipticity = 0.65)

plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent je je sais pas ce qu'il faut faire et je voudrais le savoir  printing of the R code that generated the plot nnnn. nn nn nn nn Note that the parameter was added to the code chunk to prevent printing of the R code that generated the plot. oui ou non c'est ce qu'il faut savoir 

## Et pourquoi pas ?

Eh 

#### Non 

# A nouveau au niveau 1

## On veut voir

### Si ça va

#### Aussi bien
